---
title: The Life of Eli
author: |
 | Matthew Malishev^1^*
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
output:
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 3
    number_sections: no
    toc_float: yes
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 3
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
inludes:
  before_body: 
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Data/Eli"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/Eli
classoption: portrait
# ^['https://github.com/darwinanddavis/Eli'] # footnote
vignette: >
  %\VignetteIndexEntry{Eli}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 echo = TRUE, # show code chunks in output
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\  

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview  

Activity data for Eli for his first year, including time spent feeding, sleeping, in leisure and values for growth and other behavoural traits.  

\  

TO DO  
* ~~separate activity states~~  
* separate hour and mins, then convert time to hours   

```{r eval=F, echo=F, results="hide", warnings=F}
devtools::session_info()
# !!! reinstalls all packages 
### update.packages(ask = FALSE, dependencies = c('Suggests'))

# access files from google drive
# install.packages("googledrive") # get googledrive package
# library(googledrive)
# # http://googledrive.tidyverse.org/
# # common commands: find, ls, mv, cp, mkdir, and rm
# 

# drive_find(n_max=10) # set output limits
# drive_find(type = "folder")
# drive_get("~/Data/eli/feb.csv")

```

### Load and clean data  
```{r}
setwd(params$dir) # set wd
list.files()
```
```{r}
d <- "may" # choose month or total period

data <- read.csv(paste0(d,".csv"),header=T,sep=",", stringsAsFactors=FALSE)
colnames(data) <- c("Activity","Trait","Start","Finish","Value")
data[c("Activity", "Trait")] <- sapply(data[c("Activity", "Trait")],as.character)
head(data) 

```

### Subset activities  
```{r}
unique(data$Activity)
grow <- subset(data,subset=Activity=="Growth");head(grow)
feed <- subset(data,subset=Activity=="Feeding");head(feed)
sleep <- subset(data,subset=Activity=="Sleep");head(sleep)
diaper <- subset(data,subset=Activity=="Diapering");head(diaper)
leisure <- subset(data,subset=Activity=="Leisure");head(leisure)
```

### Subset traits   
```{r}
# activity states with traits: grow,feed,diaper,leisure

# grow

# feed
breast_l <- subset(feed,subset=Trait=="Left Breast");head(breast_l)
breast_r <- subset(feed,subset=Trait=="Right Breast");head(breast_r)

# 
```

```{r echo=F, warnings=F, results="hide"}
# get time for each state by month, no need for day

#change time vectors from character to posix    
?split
?grepl
?substr
?strsplit 
?nchar

# separate into date, month, and time (by am and pm) cols. remove year. 
#opt 1
ss <- strsplit(data$Start, "-2018");ss
ss <- lapply(ss, as.character);ss
ss <- unlist(ss);ss
ss <- strsplit(ss, "-");ss
ss <- unlist(ss);ss
ss <- lapply(ss, as.character);ss

#opt 2 # gsub replaces all instances
ss <- gsub("-2018","",data$Start);head(ss)[1]
ss <- gsub("-"," ",ss);head(ss)[1]

#opt 3
month <- sub("-2018","",data$Start);month[1]
month <- strsplit(month,"-");month[1]
month <- month

?as.POSIXct(ss,tz="UTC+5") # convert to posix
month.abb
month.name
ff <- substr(ss,0,6) ;ff[1] # separate first section before period 
t <- substr(ss,9,99) ;t # separate second section after period 
t <- gsub(" pm","",t);t # remove pm
t <- gsub(" am","",t);t # remove am
# format(as.POSIXct((t) * 86400, origin = "1970-01-01", tz = "UTC"), "%H:%M") #convert to POSIX     

# add time col to df after separating (only need time difference to get activity time per activity state for plots)   
# add zero to hour
hour <- substr(t,1,2);hour# create hour col
min <- substr(t,3,99);min# create min col

#opt 4
ss <- sub("-2018","",data$Start);ss[1]
ss <- sub("-","/",ss);ss
# as.POSIXct(ss,tz="UTC+5") # convert to posix

?strptime
z <- as.difftime(c("3:20", "23:15"), format = "%H:%M") # 3rd gives NA
as.numeric(z, units = "mins")
?difftime

class(strsplit(data$Start, "-")[[101]]) # this is a character vec

#change days and months to two chars
# data$Day[nchar(data$Day)==1]<-paste(0,data$Day[nchar(data$Day)==1],sep="")
# data$Month[nchar(data$Month)==1]<-paste(0,data$Month[nchar(data$Month)==1],sep="")

#potential to separate by am and pm
strsplit(data$Start, "p")
```

```{r echo=F, eval=F, results="hide", warning=F}
#### Lab  
micro_shd<-subset(micro_shd_all, format(as.POSIXlt(micro_shd_all$dates), "%y/%m/%d")>=daystart & format(as.POSIXlt(micro_shd_all$dates), "%y/%m/%d")<=dayfin)

data$Start <- as.POSIXct(as.character(data$Trait),format="%m/%d/%Y %H:%M")
head(data)
```



