---
title: The Life of Eli
author: |
 | Matthew Malishev^1^*
 |
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
output:
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 5
    fig_height: 4
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
inludes:
  before_body: 
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Data/Eli"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/Eli
classoption: portrait
# ^['https://github.com/darwinanddavis/Eli'] # footnote
vignette: >
  %\VignetteIndexEntry{Eli}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 echo = TRUE, # show code chunks in output
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\  

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`

\newpage  

## Overview  

Activity data for Eli for his first year, including time spent feeding, sleeping, in leisure and values for growth and other behavoural traits.  

\  

**TO DO**  
* ~~separate activity states~~  
* separate hour and mins, then convert time to hours   

```{r eval=F, echo=F, results="hide", warnings=F}
devtools::session_info()
# !!! reinstalls all packages 
### update.packages(ask = FALSE, dependencies = c('Suggests'))

# access files from google drive
# install.packages("googledrive") # get googledrive package
# library(googledrive)
# # http://googledrive.tidyverse.org/
# # common commands: find, ls, mv, cp, mkdir, and rm
# 

# drive_find(n_max=10) # set output limits
# drive_find(type = "folder")
# drive_get("~/Data/eli/feb.csv")

```

### Install dependencies
```{r, load packages, include=T, cache=F, message=F, warning=F, results='hide'}
packages <- c("stringi","tidyr","sp","RColorBrewer","ggplot2","ggthemes","RCurl")   
if (require(packages)) {
    install.packages(packages,dependencies = T)
    require(packages)
}
ppp <- lapply(packages,require,character.only=T)
if(any(ppp==F)){cbind(packages,ppp);cat("\n\n\n ---> Check packages are loaded properly <--- \n\n\n")}
```

### Set plotting function  
```{r results="hide", echo=F}

script <- getURL("https://raw.githubusercontent.com/darwinanddavis/plot_it/master/plot_it.R", ssl.verifypeer = FALSE)
eval(parse(text = script))

cat("plot_it( \n0 for presentation, 1 for manuscript, \nset colour for background, \nset colour palette. use 'display.brewer.all()', \nset alpha for colour transperancy, \nset font style \n)")
plot_it(0,"blue","Spectral","Greens",1,"mono") # set col function params
plot_it_gg("white") # same as above 
```

### Load and clean data  
```{r}
setwd(params$dir) # set wd
list.files()
```

```{r}
d <- "may" # choose month or total period

data <- read.csv(paste0(d,".csv"),header=T,sep=",", stringsAsFactors=FALSE)
colnames(data) <- c("Activity","Trait","Start","Finish","Value")
data[c("Activity", "Trait")] <- sapply(data[c("Activity", "Trait")],as.character)
head(data) 

# load june july aug + data
# skip first three redundant rows
dd <- "junejulyaug"
data2 <- read.csv(paste0(dd,".csv"),header=T,sep=",", stringsAsFactors=FALSE,skip=3)
colnames(data2) <- c("Activity","Trait","Start","Finish","Value")
data2[c("Activity", "Trait")] <- sapply(data2[c("Activity", "Trait")],as.character)

# add june july aug + data to existing data frame
data <- rbind(data,data2)
head(data)
```

### Subset activities  
```{r}
unique(data$Activity)
grow <- subset(data,subset=Activity=="Growth");head(grow)
feed <- subset(data,subset=Activity=="Feeding");head(feed)
sleep <- subset(data,subset=Activity=="Sleep");head(sleep)
diaper <- subset(data,subset=Activity=="Diapering");head(diaper)
leisure <- subset(data,subset=Activity=="Leisure");head(leisure)
```

### Subset traits   
```{r}
# activity states with traits: grow,feed,diaper,leisure

# grow
grow <- within(grow, rm("Finish")) # only time stamp, so remove Finish time col
head <- subset(grow,subset=Trait=="Head");head
height <- subset(grow,subset=Trait=="Height");height
weight <- subset(grow,subset=Trait=="Weight");weight  

# feed
feed <- within(feed,rm("Value")) # no values, so remove Values col
breast_l <- subset(feed,subset=Trait=="Left Breast");head(breast_l)
breast_r <- subset(feed,subset=Trait=="Right Breast");head(breast_r)
bottle <- subset(feed,subset=Trait=="Bottle");head(bottle)

# diaper
diaper <- within(diaper, rm("Finish")) # only time stamp, so remove Finish time col
pee <- subset(diaper,subset=Trait=="Pee");head(pee)
poo <- subset(diaper,subset=Trait=="Poo");head(poo) 
both <- subset(diaper,subset=Trait==unique(diaper$Trait)[1]);head(both) 

# leisure  
leisure <- within(leisure, rm("Value")) # no values, so remove Values col
bath <- subset(leisure,subset=Trait=="Bath time");head(bath)
tummy <- subset(leisure,subset=Trait=="Tummy time");head(tummy)
outdoors <- subset(leisure,subset=Trait=="Outdoors");outdoors
play <- subset(leisure,subset=Trait=="Play time");head(play)

# sleep
sleep <- within(sleep,rm("Value")) # no values, so remove Values col
sleep <- within(sleep,rm("Trait")) # no Traits, so remove Traits col

```

```{r echo=F, warnings=F, results="hide", eval=F}
# get time for each state by month, no need for day

# separate day month and years
# from heatmap ref in 'figures' Presentations Github 
require(dplyr); require(tidyr); require(Interpol.T);require(lubridate)
data<- data(Trentino_hourly_T,package = "Interpol.T")
names(h_d_t)[1:5]<- c("stationid","date","hour","temp","flag")
df<- tbl_df(h_d_t) %>%
  filter(stationid =="T0001")

df<- df %>% mutate(year = year(date),
                   month = month(date, label=TRUE),
                   day = day(date))


# current method (17-9-18)
ss <- strsplit(data$Start, "-2018");head(ss)
gsub("[^[:digit:]]", "", ss)[1] # get just numbers
gsub("[^[:digit:]]", " ", ss)[1] # get just numbers
gsub("[[:digit:]]", "", ss)[1] # get just characters

# testing how to get difference in time cols 
test <- breast_l[3,]
test[,c(3,4)]
test$Start
as.Date("18/02/18", "%d/%m/%y") # testing as.Data function
test$diff_in_days<- difftime(test$Start ,test$Finish , units = c("hours"))
as.Date("10:45 pm", "%m/%d/%y")

```

```{r}

#change time vectors from character to posix    
?split
?grepl
?substr
?strsplit 
?nchar
?stri_sub # stringi package

# separate into date, month, and time (by am and pm) cols. remove year. 
#opt 1
ss <- strsplit(data$Start, "-2018");head(ss)
ss <- lapply(ss, as.character);head(ss)
ss <- unlist(ss);head(ss)
ss <- strsplit(ss, "-");head(ss)
ss <- unlist(ss);head(ss)
ss <- lapply(ss, as.character);head(ss)


```


```{r echo=F, warnings=F, results="hide"}
#opt 2 # gsub replaces all instances
ss <- gsub("-2018","",data$Start);head(ss)[1]
ss <- gsub("-"," ",ss);head(ss)[1]

#opt 3
month <- sub("-2018","",data$Start);month[1]
month <- strsplit(month,"-");month[1]
month <- month

?as.POSIXct(ss,tz="UTC+5") # convert to posix
month.abb
month.name
ff <- substr(ss,0,6) ;ff[1] # separate first section before period 
t <- substr(ss,9,99) ;head(t) # separate second section after period 
t <- gsub(" pm","",t);head(t) # remove pm
t <- gsub(" am","",t);head(t) # remove am
# format(as.POSIXct((t) * 86400, origin = "1970-01-01", tz = "UTC"), "%H:%M") #convert to POSIX     

# add time col to df after separating (only need time difference to get activity time per activity state for plots)   
# add zero to hour
hour <- substr(t,1,2);hour[1]# create hour col
min <- substr(t,3,99);min[1]# create min col

#opt 4
ss <- sub("-2018","",data$Start);ss[1]
ss <- sub("-","/",ss);head(ss)
# as.POSIXct(ss,tz="UTC+5") # convert to posix

?strptime
z <- as.difftime(c("3:20", "23:15"), format = "%H:%M") # 3rd gives NA
as.numeric(z, units = "mins")

class(strsplit(data$Start, "-")[[101]]) # this is a character vec

#change days and months to two chars
# data$Day[nchar(data$Day)==1]<-paste(0,data$Day[nchar(data$Day)==1],sep="")
# data$Month[nchar(data$Month)==1]<-paste(0,data$Month[nchar(data$Month)==1],sep="")

#potential to separate by am and pm
ap <- strsplit(data$Start, "p"); head(ap)
```

### Growth  
No time period, just values  

#### Head  
```{r}
require("stringi")
require("tidyr")
require("sp")
require("RColorBrewer")

plot_it(0,"blue","Blues","Greens",0.5,"HersheySans")
hv <- gsub("[^[:digit:]]", "", head$Value) # get just integers
stri_sub(hv,3,1) <- ".";hv # insert the decimal point in the correct place
head$Value <- hv %>% as.numeric() # make numeric

d <- head
par(las=1,bty="n")
ylim <- round(max(d$Value,10))
with(d,plot(Value,
            col=colv,
            type="b",lwd=3,
            pch=20,
            ylim=c(0,ylim),
            ylab="Head circumference (cm)",
            xlab="Time",
            xaxt="n"
))
axis(1,at=1:4,labels=month.abb[2:5])
# started solids July 2 abline()
title("Head circumference (cm)")
```

#### Weight   
```{r}
require(RColorBrewer)
require(sp)

plot_it(0,"blue","Blues","Greens",0.5,"HersheySans")
wv <- gsub("[^[:digit:]]", "", weight$Value) # get just integers
stri_sub(wv,2,1) <- ".";wv # insert the decimal point in the correct place
weight$Value <- wv %>% as.numeric() # make numeric

d <- weight
par(las=1,bty="n")
xlim <- length(weight$Value)
ylim <- round(max(weight$Value,10))
with(d,plot(Value,
            col=colv,
            type="b",lwd=3,
            pch=20,
            ylim=c(0,ylim),
            ylab="Weight (kg)",
            xlab="Time",
            xaxt="n"
))
# axis(1,at=1:xlim,labels=rep(month.abb[2:(xlim/2)],each=2))
# started solids July 2 abline()
title("Weight (kg) over time")
```

#### Height  
```{r}
require(RColorBrewer)
require(sp)

plot_it(0,"blue","Blues","Greens",0.5,"HersheySans")
hhv <- gsub("[^[:digit:]]", "", height$Value) # get just integers
stri_sub(hhv,3,1) <- "." ;hhv # insert the decimal point in the correct place
height$Value <- hhv %>% as.numeric() # make numeric

d <- height
par(las=1,bty="n")
with(d,plot(Value,
            col=colv,
            pch=20,
            type="b",lwd=3,
            ylim=c(0,70),
            ylab="Height (cm)",
            xlab="Time",
            xaxt="n"
))
axis(1,at=1:4,labels=month.abb[2:5])
# axis(1,at=c(0,length(d$Value)),labels=c("",""))# bookending axis tick marks
# started solids July 2 abline()
title("Height (cm) over time")

```

### Feeding  
Only time period, no values 
Started solids July 2      

#### Left breast  
Only time period, no values    
```{r}
# started solids July 2    
head(breast_l)
```

#### Right breast  
Only time period, no values      
```{r}
# started solids July 2    
head(breast_r)
```


### Diaper  
Only time stamp (count)  
Started solids July 2    
```{r}
# started solids july 2  
unique(diaper$Trait)
```

#### Pee  
Only time stamp (count)  
```{r}
# started solids july 2  
which(pee$Start!=pee$Finish) # are there any time periods > 1?
nrow(pee)
```

#### Poo  
Only time stamp (count)  
```{r}
# started solids july 2  
which(poo$Start!=poo$Finish) # are there any time periods > 1?
nrow(poo)
```

#### Both    
Only time stamp (count)  
```{r}
# started solids july 2  
which(both$Start!=both$Finish) # are there any time periods > 1?
nrow(both)
```

### Leisure  
Only time period, no values  
```{r}

```

#### Bath
Only time period, no values  
```{r}

```

#### Tummy  
Only time period, no values  
```{r}

```

#### Outdoors
Only time period, no values  
```{r}

```

#### Play  
Only time period, no values  
```{r}

```

### Sleep  
No values, just time period  
```{r}
head(sleep) 
```

### Lab  
```{r echo=F, eval=F, results="hide", warning=F}
#### Lab  
micro_shd<-subset(micro_shd_all, format(as.POSIXlt(micro_shd_all$dates), "%y/%m/%d")>=daystart & format(as.POSIXlt(micro_shd_all$dates), "%y/%m/%d")<=dayfin)

data$Start <- as.POSIXct(as.character(data$Trait),format="%m/%d/%Y %H:%M")
head(data)
```



